Distribution-oblivious Programming with Dish

### 2019-11-21 pldi/draft.tex
A draft on the distribution planner, graph transformations

Can we find a solution for the commands in coreutils?

@@

@@

In order to effectively apply optimizations and distribute a shell
script, it first has to be translated to a manipulable
representation. As mentioned in \ref{}, the two main abstractions of
the shell, are files that contain data, and commands that communicate
through these files. We propose a slight variation of the standard
dataflow graph, that is commonly found in stream processing systems,
as the representation of a distributed shell computation. The main
benefit of the dataflow graph model is that it clearly exposes
operator level parallelism, as different nodes can be independent
processing workers. In our case each node represents a command, and
each edge represents a file.

@@

\kk{Partially checked Assumption. The same file exists once in the
  substree. This is a conservative assumption that ensures that no two
  commands write or read to the same file concurrently. We kind of
  check it but not completely. We check that the expanded evaluated
  files are not the same, but one could have part of a file in a
  variable etc.}

@@

\kk{Mention that different nodes can be safely executed independently,
  and that this exposes operator level parallelism}

\subsection{Scaling up the distributed computation/Vertically expanding the graph/Graph Transformations/ Equations}

The planner acts on the dataflow graph, by first expanding it ``as
much as possible'' to expose parallelization opportunities. It uses
knowledge about the inputs of each stateless and pure command.

\kk{Explain the algorithm and how it pushes parallelization through
  the graph, while preserving correctness.}

\kk{Give an example}

\kk{Can we have a semi-formal argument here?}

\kk{Explain how our graph transformations expose data parallelism as
  operator level parallelism.}

\kk{Different transformations to explain: Stateless, Pure, File
  splits, XARGS. Mention as much as possible formally using the
  definition of a node.}

The following equation holds for any stateless command $f$.

\[
f(x.x') = f(x).f(x')
\]

This is the basis of all optimizations that the graph does, and allows
for correct data parallelism.

@@

@@

@@

\section{Implementation}

@@

@@

>>>>

\kk{Can we find a solution for the commands in coreutils?}

\kk{Important TODO: Find at least one more command that can be
  executed like sort. Does wc work? If there is no pure command that
  can be easily parallelized other than sort, this story is not very
  nice.}

@@

\kk{Inline listing?}

@@

In order to effectively apply optimizations and distribute a shell
script, it first has to be translated to a manipulable
representation. As mentioned in \ref{}, the two main abstractions of
the shell, are files that contain data, and commands that communicate
through these files. We propose a slight variation of the standard
dataflow graph, that is commonly found in stream processing systems,
as the representation of a distributed shell computation. The main
benefit of the dataflow graph model is that it clearly exposes
operator level parallelism \kk{Is this a common term? Is there any
  definitive citation for it?}, as different nodes can be independent
processing workers. In our case each node represents a command, and
each edge represents a file.

@@

\kk{Partially checked Assumption. Only one reader and one writer to
  each file, and no cycles. OLD MORE CONSERVATIVE ASSUMPTION: The same
  file exists once in the substree. This is a conservative assumption
  that ensures that no two commands write or read to the same file
  concurrently. We kind of check it but not completely. We check that
  the expanded evaluated files are not the same, but one could have
  part of a file in a variable etc.}

@@

\kk{I think that there is not enough content for the planner section
  and it should probably just be called graph transformations. Then we
  could have a small subsection at the end about how the planner
  applies transformations iteratively starting from the source of the
  graph and pushing them until the end.}

As mentioned in \Cref{ir}, the dataflow graph exposes operator level
parallelism. Different nodes in the graph can be executed
concurrently, communicating through the edges. Operator level
parallelism is already exposed by the shell, by using pipes \texttt{|}
and the background operator \texttt{\&}. Unfortunately, simple
operator level parallelism leaves a lot to be desired, since scaling a
computation requires manually introducing more operators.

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. Data level parallelism is the
basic optimization that is used by popular distributed processing
systems such as MapReduce\kk{cite} and Apache Spark\kk{cite}.

\subsection{Graph transformations}

In our case, this is achieved by iteratively applying graph
transformations that preserve correctness, while exposing
parallelization opportunities. We illustrate that with a motivating
example; consider the following program and its dataflow graph.

\begin{lstlisting}[language=sh, float=h, numbers=none, escapeinside={($}{$)}]
 grep ``foo'' f1 f2 // Add its graph
\end{lstlisting}

\noindent
In this example, \texttt{grep} is applied to files f1 and f2, and its
result is sent to standard output. However, the result would be the
same if we concatenated the results of \texttt{grep ``foo'' f1} and
\texttt{grep ``foo'' f2}. This intuition can be generalized to all
stateless commands, as they are all essentially ``linear'' with
respect to concatenation and the following equation holds.

\[
\forall f, \mathrm{stateless}(f) \implies f(x.x') = f(x).f(x')
\]

\kk{Complete this story with pure, after we have written the section
  about command categories.}

\subsection{Planner Algorithm}

The \texttt{Dish} planner extends the transformations that were
described above with edge splits in order to allow for the above
transformations to be performed. More precisely, it splits input files
and graph edges into several chunks so that they can be processed in
parallel by the next levels.

\kk{Explain the greedy algorithm that our planner performs. It splits
  input files, and then pushes parallelization through the graph,
  starting from the source nodes.}

\kk{Maybe give an example}

@@

\kk{I think we should actually remove this completely.}

@@

\kk{This whole section for just in itme planning, is also relevant for
  the front end. If the front end is executed just in time, it would
  have more information, and maybe it could make a sound analysis.}

@@

\section{Evaluation}

@@

\kk{We should have at least some pipelines that have a bad pure
  command in them.}

@@

\kk{We need to cite DGSH}
